---
title: "Pirate_OS_ET_Analysis"
output:
  pdf_document: default
  html_document:
    fig_height: 8
    fig_width: 12
editor_options:
  chunk_output_type: console
---
# Load Packages and Data
```{r Load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(vroom)
library(remotes)
library(plotrix)
library(ggpattern)

```

```{r Load participant responses, message=FALSE, warning=FALSE, results='hide'}
# Participant responses are stored in separate text files
# Load data from these files into one data frame
list_of_files <- list.files(path = "./data", recursive = FALSE,
                            pattern = "\\.txt$", 
                            full.names = TRUE)
dataP_raw <- vroom(list_of_files, id = "Code")
dataP_raw<-dataP_raw%>%
  separate(Code,c("path","path2","Code"), sep="/")

# Select only columns necessary for analysis
dataP<-dataP_raw%>%
  select("Code",
         "Begin Time - msec",
         "End Time - msec",
         "Duration - msec",
          "Fixations",
         "IP",
         "Trial",
        "Behavioral")
dataP$Code<-str_replace_all(dataP$Code, ".txt", "")

```

```{r Load trial and list data}
# Trial data identifies the four locations of interest for each trial
TrialdataP<-read_tsv("./data/Suppdata/Trial_info_corrected.txt", col_names =TRUE)
head(TrialdataP)

# List data identifies the trial version each participant took part in
List<-read_csv("./data/Suppdata/List.csv", col_names =TRUE)
head(List)

```

# Data Processing
```{r Fix behavioral data}
# Filter out NA's in the 'IP' and 'Fixations' columns
dataP<-dataP%>%
  filter(!is.na(IP))%>%
  filter(!is.na(Fixations))

# Get unique values for behavioral data for each trial and participant
Behave<-dataP%>%
  select(Code, Trial, Behavioral)%>%
  filter(!is.na(Behavioral))%>%
  unique()%>%
  mutate(Behavioral = factor(Behavioral))

# Left join behavioral data to main data frame
dataP<-dataP%>%
  select("Code",
         "Begin Time - msec",
         "End Time - msec",
         "Duration - msec",
          "Fixations",
         "IP",
         "Trial")%>%
  left_join(Behave)%>%
  mutate(Behavioral = factor(Behavioral),
           IP = factor(IP),
           Trial = factor(Trial),
           Code = factor(Code))%>%
  rename(IP_start = `Begin Time - msec`,
         IP_end = `End Time - msec` ,
         Duration = `Duration - msec` 
         )%>%
  separate(Trial, c("Item","Experiment", "Condition"), remove=FALSE)%>%
  # Filter for 'Pirate' experiment, as the dataP currently includes data for
  # a related experiment that was conducted simultaneously
  filter(Experiment == "Pirate")

```

```{r Join trial info and list info}
dataP<-dataP%>%
  left_join(List)%>%
  unite(ListCode, c("Trial", "List"), sep="_",remove = FALSE)%>%
  # Here we use the trial code and list info to join in the trial info,
  # which let us know what participants were looking at in each trial - this is
  # known as the 'Fixation Object'
  left_join(TrialdataP, by = c("ListCode" = "ListCode"))%>%
  select( "Code",
          "Duration",
          "Fixations",
          "TargLoc",
          "CompLoc",
          "Dist1Loc",
          "Dist2Loc",
          "IP",
          "Trial",
          "Item",
          "Behavioral",
           "ListCode")

```

```{r Code fixation object}
dataP$FixationObject<-ifelse(dataP$Fixations==dataP$TargLoc,"Target",
                             ifelse(dataP$Fixations==dataP$CompLoc,"Competitor",                                                
                                    ifelse(dataP$Fixations==dataP$Dist1Loc,"Dist1","Dist2")
                                    )
                            )
# Remove unnecessary columns
dataP<-dataP%>%
  select( "Code",
          "Duration",
          "Fixations",
          "TargLoc",
          "CompLoc",
          "Dist1Loc",
          "Dist2Loc",
          "IP",
          "Trial",
          "Item",
          "Behavioral",
          "FixationObject")
                            
```

```{r Create columns for fixation proportions}
# For each time interval, the location the participant is looking at is coded
# as '1' and the others are coded as '0'
dataP$TargetCount<-ifelse(dataP$FixationObject=="Target",dataP$Duration,0)
dataP$CompetitorCount<-ifelse(dataP$FixationObject=="Competitor",dataP$Duration,0)
dataP$Dist1Count<-ifelse(dataP$FixationObject=="Dist1",dataP$Duration,0)
dataP$Dist2Count<-ifelse(dataP$FixationObject=="Dist2",dataP$Duration,0)

# Remove unnecessary columns
dataP<-dataP%>%
  select( "Code",
          "Duration",
          "Fixations",
          "TargLoc",
          "CompLoc",
          "Dist1Loc",
          "Dist2Loc",
          "IP",
          "Trial",
          "Item",
          "Behavioral",
          "FixationObject",
          "TargetCount",
          "CompetitorCount",
          "Dist1Count",
          "Dist2Count"
        )

```

```{r Calculate fixation proportions}
# Sum fixation totals grouped by trial
(dataP_prop<-dataP%>%
  group_by(Code,Trial,IP)%>%
   summarize(TargetSum=sum(TargetCount),
             CompetitorSum=sum(CompetitorCount),
             Dist1Sum=sum(Dist1Count),
             Dist2Sum =sum(Dist2Count),
             IPtotal = sum(TargetSum,CompetitorSum,Dist1Sum,Dist2Sum)
            )
 )

# Calculate fixation proportions
dataP_prop<-dataP_prop%>%
  mutate("Target (chest)" = TargetSum/IPtotal,
          "Verb Related (bone)" = CompetitorSum/IPtotal,
          "Agent Related (ship)" = Dist1Sum/IPtotal,
          "Unrelated (cat)" = Dist2Sum/IPtotal
          )%>%
  select(Code,
        Trial,
        IP,
        "Target (chest)",
        "Verb Related (bone)",
        "Agent Related (ship)",
        "Unrelated (cat)",
        IPtotal
        )%>%
  gather("Target (chest)",
        "Verb Related (bone)",
        "Agent Related (ship)",
        "Unrelated (cat)",
        key = "Fixation Object",
        value = "Fixation_Proportion")

```

# Data Visualization
```{r Prepare data for analysis}
# Ensure factor variables are properly encoded as such
dataP_prop2<-dataP_prop%>%
  ungroup()%>%
  separate(Trial, c("Item","Experiment","Condition"), remove = FALSE)%>%
  mutate(`Fixation Object` = factor(`Fixation Object`, levels = c("Target (chest)","Verb Related (bone)","Agent Related (ship)","Unrelated (cat)")),
          Item = factor(Item) )%>%
  # Re-code interest periods for clarity
  mutate(IP=recode(IP, 
                    "np1"="Noun 1",
                    "verb"="Verb",
                    "np2"="Noun 2"),
         IP=factor(IP, levels=c("Noun 1", "Verb", "Noun 2")))

# Calculate mean and standard error of fixation proportions for each interest
# period, fixation object, condition, and experiment
(dataP_prop_summary<-dataP_prop2%>%
 group_by(IP,`Fixation Object`,Condition,Experiment)%>%
   summarize(`Fixation Proportion Mean` = mean(Fixation_Proportion,na.rm=TRUE),
               Fixation_ProportionSE = std.error(Fixation_Proportion,na.rm=TRUE)
            )
)

droplevels(dataP_prop_summary)

```

```{r Main plot}
# The palette with gray:
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# The palette with black:
cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Plot mean fixation proportion by item across the three interest periods
Pirate_ByIP_Base<-ggplot(data=dataP_prop_summary,
                         aes(`Fixation Object`, `Fixation Proportion Mean`)
                         )+
  facet_wrap(~IP)+
  geom_col(aes(fill=`Fixation Object`)
           )+
  theme(legend.title = element_blank(),
        legend.key.size = unit(1, 'cm'),
        text = element_text(family="serif",size=25),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()
        )+
  scale_fill_manual(values=cbbPalette)+
  geom_errorbar(aes(ymin = `Fixation Proportion Mean`-Fixation_ProportionSE,
                    ymax=`Fixation Proportion Mean`+Fixation_ProportionSE,
                    group = Condition), 
                width=.2,position=position_dodge(0.9)
                )+
  coord_cartesian(ylim = c(0.0, 1))

```

## Analysis of Main Plot

Fixation proportions are random during the Noun 1 interest period, as
participants have not yet received the verbal cue to predict the target noun.
The elevated proportion of fixations to the target noun compared to all others 
uggests that participants were able to predict the target noun. Participants
hear the target noun during the Noun 2 interest period; thus, the vast majority
of fixations are directed towards the target.
